---
title: "Validation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

validation schema for 2000 oligo pool.
- 3 shortened bacterial sequences
- (240 total) from the ACGT random oligos, picked out 20 sequences that had neither CA, AAUAAA, UGUA, or GUGUG and then added back motifs
  - null 
  - null + CA
  - null + UGUA 
  - null + AAUAAA
  - null + GUGUGUUGU
  - null + CA + AAUAAA
  - null + CA + UGUA
  - null + CA + GUGUGUUGU
  - null + CA + AAUAAA + UGUA
  - null + CA + AAUAAA + GUGUGUUGU
  - null + CA + UGUA + GUGUGUUGU
  - null + CA + AAUAAA + UGUA + GUGUGUUGU
- (80 total) picked 20 sequences that had one strong AAUAAA site.
  - ACTCAA,ACTCAA,ACTAAA permutations 
- (80 total) picked 20 sequences that had one strong UGUA site 
  - UGCA,CGUA,UCUA permutations 
- (total 1554) picked 222 sequences (111 from each species) sequences for in-silico evolution 
  - rounds (0,3,10) in either maize,tobacco, or both 
- after removing duplicates (total is 1813), I have 187 sequences left
  - normalize each dataset enrichment 0 to 1 and then take the average of both. 
  - 62 top scorers
  - 62 mid scorers
  - 63 low scorers
  
```{r}
library(tidyverse)
library(dplyr)
library(PNWColors)
library(ggsignif)
library(stringr)
library(readr)
library(ggrepel)
library(gprofiler2)
library(Biostrings)
library(universalmotif)
library(ggpubr)
library(MetBrewer)
library(PLColors)
```


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
load('/ProtoplastSTARRseq/RData/all_terms.Rdata')
```


```{r}
all_terms_RNA <- all_terms %>% mutate(seq2=gsub("T","U",seq))
all.seqs.RNA <- Biostrings::RNAStringSet(deframe(select(all_terms_RNA, id, seq2)))
all.seqs.RNA
```

```{r}
proto_meme <- read_meme('../ProtoplastSTARRseq/data/protoplast_RNA.meme')
summarise_motifs(proto_meme)$name
```


#validation sequences
```{r}
ACGT_rand_RNA <- all_terms %>% mutate(seq2=gsub("T","U",seq)) %>% filter(Controls=="ACGT_random") 
ACGT_rand.seqs.RNA <- Biostrings::RNAStringSet(deframe(select(ACGT_rand_RNA, id, seq2)))
ACGT_rand.seqs.RNA
```


picking out sequences that dont really have UGUA 
```{r}
thresh <- 0.77
UGUA_not.pos <- as_tibble(scan_sequences(filter_motifs(proto_meme, name = "8-UGUA"), ACGT_rand.seqs.RNA, RC = FALSE, threshold = 1, nthreads = 0)) %>% mutate(
    score = (score - min.score) / (max.score - min.score)
  ) %>%
  group_by(sequence) %>%
  filter(score == max(score)) %>%
  filter(score <= thresh) %>%
  mutate(
    pos = 0.5 * (start + stop)
  ) %>%
  ungroup()
UGUA_not.pos
no_UGUA <- UGUA_not.pos %>% select(sequence,score,start) %>% mutate(no_UGUA=TRUE)
no_UGUA 
```


picking out sequenes that dont have AAUAAA
```{r}
thresh <- 0.81
AAUAAA_not.pos <- as_tibble(scan_sequences(filter_motifs(proto_meme, name = "9-AAUAAA"), ACGT_rand.seqs.RNA, RC = FALSE, threshold = 1, nthreads = 0)) %>% mutate(
    score = (score - min.score) / (max.score - min.score)
  ) %>%
  group_by(sequence) %>%
  filter(score == max(score)) %>%
  filter(score <= thresh) %>%
  mutate(
    pos = 0.5 * (start + stop)
  ) %>%
  ungroup()
#AAUAAA_not.pos
no_AAUAAA <- AAUAAA_not.pos %>% select(sequence,score,start) %>% mutate(no_AAUAAA=TRUE)
no_AAUAAA 
```


picking out sequences that are not G/U rich after the cleavage site 
```{r}
ACGT_rand_RNA_end <- all_terms %>% mutate(seq2=gsub("T","U",seq)) %>% filter(Controls=="ACGT_random") %>% mutate(seq2=str_sub(seq2,151,170))
ACGT_rand.seqs.RNA_end <- Biostrings::RNAStringSet(deframe(select(ACGT_rand_RNA, id, seq2)))
```


```{r}
thresh <- 0.77
GUGUG_not.pos <- as_tibble(scan_sequences(filter_motifs(proto_meme, name = "3-GUGUGUUGU"), ACGT_rand.seqs.RNA_end, RC = FALSE, threshold = 1, nthreads = 0)) %>% mutate(
    score = (score - min.score) / (max.score - min.score)
  ) %>%
  group_by(sequence) %>%
  filter(score == max(score)) %>%
  filter(score <= thresh) %>%
  mutate(
    pos = 0.5 * (start + stop)
  ) %>%
  ungroup()
GUGUG_not.pos
no_GUGUG <- GUGUG_not.pos %>% select(sequence,score,start) %>% mutate(no_GUGUG=TRUE)
no_GUGUG
```


combining all that have neither of the 3 motifs 
```{r}
null_motifs <- no_UGUA %>% 
  inner_join(no_AAUAAA,by="sequence") %>%
  inner_join(no_GUGUG,by="sequence") %>%
  mutate(average_score=(score.y+score.x+ score)/3) %>%
  left_join(all_terms %>% mutate(sequence=id) %>% select(sequence,Species,seq),by="sequence") %>%
  distinct(sequence,Species,seq,average_score) %>% 
  arrange(average_score) %>%
  mutate(has_CA = if_else(substr(seq,150,151)=="CA",TRUE,FALSE)) %>% 
  filter(has_CA == FALSE) %>% 
  head(n=20) %>%
  select(-has_CA)
null_motifs
```

all combinations
null (regular)

each additions: 
null + CA
null + UGUA 
null + AAUAAA
null + GUGUGUUGU

single additions with cleavage
null + CA + AAUAAA
null + CA + UGUA
null + CA + GUGUGUUGU

double additions with cleavage
null + CA + AAUAAA + UGUA
null + CA + AAUAAA + GUGUGUUGU
null + CA + UGUA + GUGUGUUGU

triple addition with cleavage
null + CA + AAUAAA + UGUA + GUGUGUUGU

#make dataframe with all singles 
```{r}
null_motifs_single <- null_motifs %>% 
  mutate(null=seq) %>%
  mutate(add_CA = `str_sub<-`(seq, 150, 151, value = 'CA')) %>%
  mutate(add_UGUA = `str_sub<-`(seq, 108, 111, value = 'TGTA')) %>%
  mutate(add_AAUAAA = `str_sub<-`(seq, 127, 132, value = 'AATAAA')) %>%
  mutate(add_GUGUG = `str_sub<-`(seq, 159, 167, value = 'GTGTGTTGT')) 
null_motifs_single
```


```{r}
null_motifs_all <- null_motifs_single %>% 
  #single additions with cleavage
  mutate(CA_AAUAAA = `str_sub<-`(add_CA, 127, 132, value = 'AATAAA')) %>%
  mutate(CA_UGUA= `str_sub<-`(add_CA, 108, 111, value = 'TGTA')) %>%
  mutate(CA_GUGUG= `str_sub<-`(add_CA, 159, 167, value = 'GTGTGTTGT')) %>%
  #double additions with cleavage
  mutate(CA_AAUAAA_UGUA = `str_sub<-`(CA_AAUAAA, 108, 111, value = 'TGTA')) %>%
  mutate(CA_AAUAAA_GUGUG = `str_sub<-`(CA_AAUAAA, 159, 167, value = 'GTGTGTTGT')) %>%
  mutate(CA_UGUA_GUGUG = `str_sub<-`(CA_UGUA, 159, 167, value = 'GTGTGTTGT')) %>%
  #triple addition with cleavage
  mutate(CA_AAUAAA_UGUA_GUGUG = `str_sub<-`(CA_AAUAAA_UGUA, 159, 167, value = 'GTGTGTTGT')) %>%
  select(-seq)

```

```{r}
null_motifs_final <- null_motifs_all %>% 
  pivot_longer(!c(sequence,average_score,Species),names_to="Type",values_to="final_seq") %>%
  select(sequence,Type,final_seq) %>%
  mutate(id=paste(sequence,Type,sep="|")) %>%
  select(id,final_seq) %>%
  mutate(Val_group = "null_plus_motifs")
null_motifs_final #%>% mutate(length=nchar(final_seq)) %>% ungroup() %>% count(length)
```

#count duplicates
```{r}
null_motifs_all %>% 
  pivot_longer(!c(sequence,average_score,Species),names_to="Type",values_to="final_seq") %>%
  group_by(final_seq) %>%
  filter(n()>1) %>%
  ungroup() #%>% 
  #distinct(final_seq, .keep_all=TRUE)
```


#finding sequences that have a single strong AAUAAA site. Only pick the top 20. 
```{r}
thresh <- 0.4
high_DAAUAAAR.pos <- as_tibble(scan_sequences(filter_motifs(proto_meme, name = "9-AAUAAA"), all.seqs.RNA, RC = FALSE, threshold = 1, nthreads = 0)) %>% mutate(
    score = (score - min.score) / (max.score - min.score)
  ) %>%
  filter(score >= thresh) %>%
  group_by(sequence) %>%
  arrange(desc(score)) %>%
  slice_max(score,n=2) %>%
  mutate(diff=max(score)-min(score)) %>%
  ungroup() %>% 
  select(motif,sequence,diff,score,match,start,stop) %>%
  arrange(desc(diff)) %>%
  filter(match=="AAUAAA" & diff > .35 & start > 100) %>% 
  head(n=20)
#high_DAAUAAAR.pos
```
#the permutations to add
ACTCAA,ACTCAA,ACTAAA

```{r}
final_AAUAAA_perms <- high_DAAUAAAR.pos %>% select(sequence,start,stop) %>% 
  mutate(id=sequence) %>% 
  left_join(all_terms %>% select(id,seq),by="id") %>% 
  select(id,start,stop,seq) %>%
  mutate(AATAAA=seq) %>%
  mutate(ACTAAA=`str_sub<-`(seq, start, stop, value = 'ACTAAA')) %>%
  mutate(AATCAA=`str_sub<-`(seq, start, stop, value = 'AATCAA')) %>%
  mutate(ACTCAA=`str_sub<-`(seq, start, stop, value = 'ACTCAA')) %>%
  select(-seq,-start,-stop) %>%
  pivot_longer(!id,names_to="type",values_to="final_seq") %>% 
  mutate(id=paste(id,type,sep="|")) %>%
  select(id,final_seq) %>%
  mutate(Val_group = "AAUAAA_perm")
final_AAUAAA_perms
```


#same for UGUA 
```{r}
thresh <- 0.4
high_UGUA.pos <- as_tibble(scan_sequences(filter_motifs(proto_meme, name = "8-UGUA"), all.seqs.RNA, RC = FALSE, threshold = 1, nthreads = 0)) %>% mutate(
    score = (score - min.score) / (max.score - min.score)
  ) %>%
  filter(score >= thresh) %>%
  group_by(sequence) %>%
  arrange(desc(score)) %>%
  slice_max(score,n=2) %>%
  mutate(diff=max(score)-min(score)) %>%
  ungroup() %>% 
  select(motif,sequence,diff,score,match,start,stop) %>%
  arrange(desc(diff)) %>%
  filter(match=="UGUA" & diff > .24 & start > 90) %>% 
  head(n=20)
#high_UGUA.pos
```

UGCA,CGUA,UCUA
```{r}
final_UGUA_perms <- high_UGUA.pos %>% select(sequence,start,stop) %>% 
  mutate(id=sequence) %>% 
  left_join(all_terms %>% select(id,seq),by="id") %>% 
  select(id,start,stop,seq) %>%
  mutate(UGUA=seq) %>%
  mutate(UCUA=`str_sub<-`(seq, start, stop, value = 'TCTA')) %>%
  mutate(CGUA=`str_sub<-`(seq, start, stop, value = 'CGTA')) %>%
  mutate(UGCA=`str_sub<-`(seq, start, stop, value = 'TGCA')) %>%
  select(-seq,-start,-stop) %>%
  pivot_longer(!id,names_to="type",values_to="final_seq") %>% 
  mutate(id=paste(id,type,sep="|")) %>%
  select(id,final_seq) %>% 
  mutate(Val_group = "UGUA_perm")
final_UGUA_perms
```


```{r}
Control_terms <- all_terms %>% 
  filter(Species=="Control") %>% 
  filter(Controls=="Control") %>% 
  filter(id!="tMAS") %>%
  select(id,seq) %>%
  mutate(length=nchar(seq)) %>% 
  mutate( final_seq = case_when(
    id=="t35S" ~ substr(seq,24,193),
    id=="tAg7" ~ substr(seq,1,170),
    id=="tNOS" ~ substr(seq,1,170)
  )) %>% 
  mutate(test = nchar(final_seq)) %>%
  mutate(id=paste(id,"short",sep="|")) %>%
  select(id,final_seq) %>% 
  mutate(Val_group="bact_short")
Control_terms
```


```{r}
evolution <- read.table("TobaccoSTARRseq/MachineLearning/evolution_data.tsv",header=TRUE)
```



evolution clean out duplicate sequences
```{r}
to_remove <- evolution %>%
  filter(round %in% c(0,3,10)) %>%
  group_by(sequence) %>%
  filter(n()>1) %>%
  ungroup()
to_remove
```


```{r}
temp_evolution <- evolution %>% 
  filter(round %in% c(0,3,10)) %>%
  anti_join(to_remove) %>%
  mutate(tag=paste(round,opt_for,sep="*"))  %>%
  mutate(new_id=paste(origin,tag,sep="|")) %>%
  select(id=new_id,final_seq=sequence)
temp_evolution 
```


```{r}
to_add_back <- evolution %>% 
  filter(round %in% c(0,3,10)) %>%
  group_by(sequence) %>%
  filter(n()>1) %>%
  ungroup() %>%
  mutate(tag=paste(round,opt_for,sep="*")) %>%
  group_by(sequence) %>%
  mutate(tmp=paste(tag, collapse="+")) %>% 
  mutate(new_id=paste(origin,tmp,sep="|")) %>%
  ungroup() %>%
  distinct(new_id,.keep_all=T) %>%
  select(id=new_id,final_seq=sequence)
to_add_back
```


```{r}
evolution_fin <- rbind(temp_evolution,to_add_back)
evolution_fin
evolution_fin %>% distinct(final_seq, .keep_all=T) 
```

check for duplicate sequences
```{r}
evolution_fin %>% 
  group_by(final_seq) %>%
  filter(n()>1) %>%
  ungroup()
```


#combine them all
```{r}
validation_set <- rbind(Control_terms,null_motifs_final,final_AAUAAA_perms,final_UGUA_perms,evolution_final)
validation_set 
```

```{r}
remove <- validation_set %>%
  group_by(final_seq) %>%
  filter(n()>1) %>%
  ungroup()
remove
```


```{r}
add <- validation_set %>%
  group_by(final_seq) %>%
  filter(n()>1) %>%
  mutate(new_id=paste(id,collapse="~")) %>%
  mutate(Val=paste(Val_group,collapse = "~")) %>%
  ungroup() %>%
  select(id=new_id,final_seq,Val_group=Val) %>%
  distinct(final_seq, .keep_all = T)
add
```



```{r}
val_temp <- validation_set %>%
  anti_join(remove) 

validation_final <- rbind(add,val_temp)
validation_final
```

```{r}
validation_final %>%
  group_by(final_seq) %>%
  filter(n()>1) %>%
  ungroup()
```



```{r}
names_in_final <- validation_final %>%
  separate(id,c("id","type"),"\\|") %>% distinct(id) %>% pull(id)
#names_in_final
```


```{r}
load("TobaccoSTARRseq/RData/data_seq.Rdata")
load("ProtoplastsSTARRseq/RData/data_seq_proto.Rdata")
```


```{r}
#library("scales")
range01 <- function(x){(x-min(x))/(max(x)-min(x))}
data_seq_proto$normal_p <- range01(data_seq_proto$mean_enrichment) 
data_seq_proto %>% ungroup() %>% arrange(desc(normal_p)) %>% select(id,mean_enrichment,normal_p) 
```

```{r}
data_seq$normal_t <- range01(data_seq$mean_enrichment)
data_seq %>% ungroup() %>% arrange(desc(normal_t)) %>% select(id,mean_enrichment,normal_t) 
``` 

```{r}
data_seq_proto %>% ungroup() %>% select(id,mean_enrichment,normal_p,Controls) %>% 
  left_join(data_seq %>% select(id,mean_enrichment,normal_t),by="id") %>% 
  mutate(average=(normal_t + normal_p)*.5) %>%
  arrange(desc(average)) %>%
  ggplot() + 
  geom_histogram(aes(x=normal_t),fill="blue",alpha=.5,bins = 100) +
  geom_histogram(aes(x=normal_p),fill="orange",alpha=.3,bins = 100) + 
  geom_histogram(aes(x=average),fill="green",alpha=.3,bins = 100)
```

```{r}
testing <- data_seq_proto %>% select(id,mean_enrichment,normal_p,Controls) %>% 
  left_join(data_seq %>% select(id,mean_enrichment,normal_t),by="id") %>% 
  mutate(average=(normal_t + normal_p)*.5) %>%
  filter(!is.na(average)) 

mean(testing$average)
sd(testing$average)
median(testing$average)
```

```{r}
top_scorers <- data_seq_proto %>% ungroup() %>% select(id,mean_enrichment,normal_p,Controls,seq) %>% 
  left_join(data_seq %>% select(id,mean_enrichment,normal_t),by="id") %>% 
  mutate(average=(normal_t + normal_p)*.5) %>%
  filter(!is.na(average)) %>%
  filter(Controls!="Control") %>%
  arrange(desc(average)) %>%
  filter(!id %in% names_in_final) %>%
  mutate(Val_group="high") %>%
  head(62) %>% select(id,final_seq=seq,Val_group)
top_scorers
#top_seqs <- top_scorers %>% select(id) %>% pull(id)
```


#need to add 30 strong, 30 weak, 30 intermediate of each species 

```{r}
midrange_scorers <- data_seq_proto %>% ungroup() %>% select(id,mean_enrichment,normal_p,Controls,seq) %>% 
  left_join(data_seq %>% select(id,mean_enrichment,normal_t),by="id") %>% 
  mutate(average=(normal_t + normal_p)*.5) %>%
  filter(!is.na(average)) %>%
  filter(Controls!="Control") %>%
  arrange(desc(average)) %>%
  filter(!id %in% names_in_final) %>%
  filter(average > 0.4137389 & average < 0.42) %>%
  mutate(Val_group="mid") %>%
  head(n=62) %>% 
  select(id,final_seq=seq,Val_group)
midrange_scorers
```

```{r}
low_scorers <- data_seq_proto %>% 
  ungroup() %>% 
  select(id,mean_enrichment,normal_p,Controls,seq) %>% 
  left_join(data_seq %>% select(id,mean_enrichment,normal_t),by="id") %>% 
  mutate(average=(normal_t + normal_p)*.5) %>%
  filter(!is.na(average)) %>%
  filter(Controls!="Control") %>%
  arrange(desc(average)) %>%
  filter(!id %in% names_in_final) %>%
  mutate(Val_group="low") %>%
  tail(n=63) %>% 
 select(id,final_seq=seq,Val_group)
low_scorers
```


```{r}
range_seq <- rbind(top_scorers,midrange_scorers,low_scorers) %>% mutate(id=paste(id,Val_group,sep="|"))
range_seq
```



```{r}
validation_final_final <- rbind(validation_final,range_seq)
validation_final_final %>%
  group_by(final_seq) %>%
  filter(n()>1) %>%
  ungroup()
```

```{r}
validation_final_final %>%
  mutate(length=nchar(final_seq)) %>% filter(length==170)
```

```{r}
save(validation_final_final, file = 'RData/Validation.Rdata')
```


```{r}
validation_final_final
```


```{r}
validation_final_final %>% 
  mutate(new_seq = paste0("TAAG",final_seq,"AGGT"))
```


check for RE sites, found none. 
```{r}
RE_sites <- c("GGTCTC", "CGTCTC", "GAGACC","GAGACG")
validation_final_final %>% 
  mutate(new_seq = paste0("TAAG",final_seq,"AGGT")) %>%
  mutate(RE_site = grepl(paste(RE_sites, collapse="|"), new_seq)) %>%
  count(RE_site)
```

add adapters and fix id 
```{r}
validation_final_fin <- validation_final_final %>% 
  mutate(new_seq = paste0("TAAG",final_seq,"AGGT")) %>%
  mutate(final=paste0("GCGCCGTCTCC",new_seq,"CGAGACGGTGC")) %>% 
  select(id,final) %>%
  mutate(id = gsub("^",">",id)) 
validation_final_fin %>% write_tsv("Validation_tmp.fa", col_names =FALSE)

```

awk -v OFS='\n' '{$1=$1;print}' Validation_tmp.fa > Validation_oligo_pool.fa



```{r}
validation_final_fin %>% mutate(length=nchar(final)) %>% count(length)
```



